Inferring the effective start dates of non-pharmaceutical interventions during COVID-19 outbreaks

### 2020-05-12 ms/ms.tex
revised title by UO suggestions
udpate table2 and fig1, revised prior desciption for the new normal we expect to use

% Title page
\title{Late and early effective start of non-pharmaceutical interventions inferred during COVID-19 outbreaks }

@@

For given vector $\theta$ of model parameters (\autoref{eq:theta}), we compute the expected cumulative number of reported infected individuals $\{Y_t\}_{t=1}^{T}$ for each day (\autoref{eq:Yt}). Then, since $\tilde{X}_{t-1}$ is a function of $X_1, \ldots, X_{t-1}$, we can use \autoref{eq:P(Xt|Xt-1)} to write the probability to observe the confirmed case data $\vec{X} = (X_1, \ldots, X_T)$ as 
\begin{equation} \label{eq:likelihood}
\mathbb{L}(\theta \mid \vec{X}) = P(\vec{X} \mid \theta) = P(X_1 \mid \theta) P(X_2 \mid \tilde{X_1}, \theta) \cdots P(X_T \mid \tilde{X}_{T-1}, \theta).
\end{equation}
This defines a \emph{likelihood function} $\mathbb{L}(\theta \mid \vec{X})$ for the parameter vector $\theta$ given the data $\vec{X}$.

@@

The likelihood function is defined in \autoref{eq:likelihood}.
We define the following prior distributions on the model parameters $P(\theta)$: 
\begin{equation} \label{eq:priors}
\begin{aligned} % TODO verify before submitting
Z & \sim \mathit{Uniform}(2, 5) \\
D & \sim \mathit{Uniform}(2, 5) \\
\mu & \sim \mathit{Uniform}(0.2, 1) \\
\beta & \sim \mathit{Uniform}(0.8, 1.5) \\
\lambda & \sim \mathit{Uniform}(0, 1) \\
\alpha_1, \alpha_2 & \sim \mathit{Uniform}(0.02, 1)\\
E(0) & \sim \mathit{Uniform}(0, 3000) \\
I_u(0) & \sim \mathit{Uniform}(0, 3000) \\
\tau &\sim \mathit{TruncatedNormal}(\tau^*, 5, 1, T-2),
\end{aligned}
\end{equation}
where $\mathit{TruncatedNormal}(\mu, \sigma, a, b)$ is a truncated normal distribution with mean $\mu$ and standard deviation $\sigma$ taking values between $a$ and $b$; $T$ is the number of days in the data $\vec{X}$; and $\tau^*$ is the official start of the NPI.
Most priors follow \citet{Li2020}, with the following exceptions.
$\lambda$ is used to ensure transmission rates are lower after the start of the NPIs ($\lambda < 1$).
We checked values of $\Delta t$ larger than five days and found they generally produce lower likelihood, higher DIC (see below), and unreasonable parameter estimates.
For the effective start of NPIs $\tau$ we have also tested an uninformative uniform prior $U(1,T-1)$. DIC (see below) was lower for the truncated normal prior in \hl{most countries}. More importantly, the uninformative prior could result in non-negligible posterior probability for unreasonable $\tau$ values, such as Mar~1 in the United~Kingdom. This was probably due to MCMC chains being stuck in low posterior regions of the parameter space.
We therefore decided to use the more informative truncated normal prior.

The posterior distribution of the model parameters $P(\theta \mid \vec{X})$ is then estimated using an \textit{affine-invariant ensemble sampler for Markov chain Monte Carlo} (MCMC) implemented in the \texttt{emcee} Python package~\citep{Foreman-Mackey2013}.
The maximum a posteriori

>>>>

% Title page
\title{Inferring the effective start dates of non-pharmaceutical interventions during COVID-19 outbreaks}

@@

For given vector $\theta$ of model parameters (\autoref{eq:theta}), we compute the expected cumulative number of reported infected individuals $\{Y_t\}_{t=1}^{T}$ for each day (\autoref{eq:Yt}). Then, since $\tilde{X}_{t-1}$ is a function of $X_1, \ldots, X_{t-1}$, we can use \autoref{eq:P(Xt|Xt-1)} to write the probability to observe the confirmed case data $\vec{X} = (X_1, \ldots, X_T)$ as 
\begin{equation} \label{eq:likelihood}
\mathbb{L}(\theta \mid \vec{X}) = 
P(\vec{X} \mid \theta) = 
P(X_1 \mid \theta) \cdot P(X_2 \mid \tilde{X_1}, \theta) \cdots P(X_T \mid \tilde{X}_{T-1}, \theta).
\end{equation}
This defines a \emph{likelihood function} $\mathbb{L}(\theta \mid \vec{X})$ for the parameter vector $\theta$ given the data $\vec{X}$.

@@

The likelihood function is defined in \autoref{eq:likelihood}.
The posterior distribution of the model parameters $P(\theta \mid \vec{X})$ is estimated using an \textit{affine-invariant ensemble sampler for Markov chain Monte Carlo} (MCMC) implemented in the \texttt{emcee} Python package~\citep{Foreman-Mackey2013}.

We defined the following prior distributions on the model parameters $P(\theta)$: 
\begin{equation} \label{eq:priors}
\begin{aligned} % TODO verify before submitting
Z & \sim \mathit{Uniform}(2, 5) \\
D & \sim \mathit{Uniform}(2, 5) \\
\mu & \sim \mathit{Uniform}(0.2, 1) \\
\beta & \sim \mathit{Uniform}(0.8, 1.5) \\
\lambda & \sim \mathit{Uniform}(0, 1) \\
\alpha_1, \alpha_2 & \sim \mathit{Uniform}(0.02, 1)\\
E(0) & \sim \mathit{Uniform}(0, 3000) \\
I_u(0) & \sim \mathit{Uniform}(0, 3000) \\
%\tau &\sim \mathit{TruncatedNormal}(\tau^*, 5, 1, T-2),
\tau &\sim \mathit{TruncatedNormal}\Big(\frac{\tau^*+\tau^0}{2}, \frac{\tau^*-\tau^0}{2}, 1, T-2\Big),
\end{aligned}
\end{equation}
where the prior for $\tau$ is a truncated normal distribution shaped so that the date of the first and last NPI, $\tau^0$ and $\tau^*$ (\autoref{table:table1}), are at minus and plus one standard deviation, and taking values only between 1 and  $T-2$, where $T$ is the number of days in the data $\vec{X}$.
We have also tested an uninformative uniform prior $U(1,T-2)$. DIC (see below) was lower for the truncated normal prior in \hl{most countries}. More importantly, the uninformative prior could result in non-negligible posterior probability for unreasonable $\tau$ values, such as Mar~1 in the United~Kingdom. This was probably due to MCMC chains being stuck in low posterior regions of the parameter space. We therefore decided to use the more informative truncated normal prior.
Other priors follow \citet{Li2020}, with the following exceptions.
$\lambda$ is used to ensure transmission rates are lower after the start of the NPIs ($\lambda < 1$).
We checked values of $\Delta t$ larger than five days and found they generally produce lower likelihood, higher DIC (see below), and unreasonable parameter estimates, and therefore chose $U(1,5)$ as the prior.
